\section{Introduction}


Cloud computing is an emerging technology that provides access to powerful computing resources without the need to build and maintain in-house infrastructures. It is a broad term that refers to both the services being provisioned to users from a computing infrastructure, often referred to as \textit{Software as a Service (SaaS)} and to the hardware and software required to support this provision, categorized as \textit{Platform as a Service (PaaS)} and \textit{Infrastructure as a Service (IaaS)}. It is defined in~\cite{mell2011nist} as a model for enabling ubiquitous, convenient, on-demand network access to a shared pool of configurable computing resources that can be rapidly provisioned and released with minimal management effort or service provider interaction. As users and organizations are gradually shifting from traditional computing towards cloud computing solutions, the need for evaluation of several aspects of cloud computing, such as performance and security, becomes increasingly important.

Performing experimentation however in real cloud environments is challenging since it is expensive, time costly, and not repeatable~\cite{calheiros2013emusim} and is often constrained by the rigidity of the infrastructure. A more realistic approach is the use of cloud simulation tools. Such tools have been proposed in recent years, providing experimentation platforms for evaluating different strategies for the efficient use of cloud resources. Simulation saves the expense of experimenting on real cloud infrastructures, since a comprehensive study of the whole system, with respect to proposed improvements, in real cloud resources can be extremely difficult and prohibitively expensive. Thus, cloud simulators provide simplicity, repeatability of experiments and reduced experimentation costs.

A preliminary version of this work was published in~\cite{filelis2017}.